Occurance rate of galaxy pairs across cosmic time:\\ Local Group analogs in a cosmological context

### 2024-06-06 pears.tex
deploy: 25aee7d523218253c42225dd158be0e25da52704

\documentclass[twocolumn]{aastex63}

\newcommand\aastex{AAS\TeX}
\newcommand\latex{La\TeX}
\newcommand\msun{\rm{M}_{\odot}}
\newcommand\ID{\textit{Illustris-Dark}}
\newcommand\IH{\textit{Illustris-Hydro}}
\newcommand\SUBFIND{\texttt{SUBFIND}}
\definecolor{mypink}{RGB}{232,132,161}
\newcommand{\kc}[1]{\textcolor{mypink}{\textbf{#1}} }
\newcommand{\gb}[1]{\textcolor{olive}{\textbf{#1}} }
%\newcolumntype{H}{>{\setbox0=\hbox\bgroup}c<{\egroup}@{}}
%\usepackage{subcaption}
\usepackage{nicefrac}
\usepackage{soul}
% \usepackage{amssymb}

% \graphicspath{{./}{figures/}}
% \graphicspath{{../}{figures/}}

% \shorttitle{Sample article}
% \shortauthors{Schwarz et al.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

\title{Occurance rate of galaxy pairs across cosmic time:\\ Local Group analogs in a cosmological context}
\correspondingauthor{Katie Chamberlain} 
\email{katiechambe@email.arizona.edu}

\author[0000-0001-8765-8670]{Katie Chamberlain}
\affil{Steward Observatory, University of Arizona\\
933 North Cherry Avenue\\
Tucson, AZ, 85721, USA}
\author{Gurtina Besla}
\affil{Steward Observatory, University of Arizona\\
933 North Cherry Avenue\\
Tucson, AZ, 85721, USA}
\begin{abstract}

  % \textbf{Context} 
  Dwarf galaxies are extremely common at all redshifts and often exhibit morphological disturbances that do not primarily result from galaxy mergers. As such, pair fractions, rather than morphological indicators (asymmetry, concentration) are needed to infer dwarf galaxy  merger rates from observational surveys.
  However, in contrast to massive galaxies ($5\times 10^9\leq$ $\rm M_{prim,*} \leq 1\times 10^{11} \msun$), pair fractions for dwarf galaxies ($10^8 < \rm M_* < 5 \times 10^9 \msun$) have been theoretically constrained only at low redshift.
  % \\\textbf{Aims}
  We aim to quantify the occurence rates and kinematics of isolated dwarf and massive galaxy pairs from $z=0-4$ in the Illustris and IllustrisTNG cosmological simulations.
  % \\\textbf{Methods} 
  Dwarf and massive subhalo pairs are identified in each snapshot of the Illustris \texttt{SUBFIND} catalogs. 
  % \\\textbf{Results} 
  We find that roughly 8-10\% of dwarf galaxies will have major companion (stellar mass ratio $>1:4$) at $z=0$, increasing steadily before plateauing in \ID(\IH) at 16\%(12\%) by $z\sim 2$. The occurence rate of dwarf galaxies with minor companions (stellar mass ratios between $1:10$ and $1:4$) is roughly constant at $\sim 2.5\%$ from $z=1-4$. In contrast, the occurence rate of major and minor massive galaxy pairs remains approximately constant across time, as massive galaxies have a major (minor) companion $\sim18\%$ ($\sim14\%$) of the time.
  % \\\textbf{Conclusions} 
  This suggests that the hierarchical evolution of dwarf galaxies may proceed differently than that of massive galaxies. LMC-SMC mass analogs (dwarf minor pairs) are twice as common at redshifts greater than 2, but the present day LMC-SMC system is dynamically rare at all redshifts (in terms of pair separation and velocity). The M31-M33 pair is dynamically typical of massive minor pairs at $z\approx0$, while the MW-LMC has unusually high velocities and low relative separations for a major massive pair at low redshift. Baryonic physics impacts the occurence rate of galaxy pairs and systematically changes the kinematics of dwarf pairs, suggesting observations of dwarf galaxy pairs can be used to test subgrid physics in cosmological simulations.

\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction} \label{sec:intro}
%%%%%%%%%%%%%%%%%%%%%%
Galaxy mergers are responsible for many physical processes that shape galactic evolution, including triggered star formation~\citep{mihos96, hopkins13, patton20,hani20}, increased AGN activity~\citep{hopkins08,comerford15,ellison19}, and galaxy morphology~\citep[e.g.,][]{rg17}. Dwarf galaxies  make up a majority of the galaxy population at all redshifts (REF) and are also expected to undergo interactions and merge. Such processes may be important to understand the growth and evolution of intermediate mass black holes, which are often found to be ``wandering" in low mass galaxy hosts~\citep{stemo20,reines20}. Despite the ubiquity of dwarf galaxies in the universe, we do not have a clear path to constrain the merger rates of these systems observationally or to place known interacting dwarf galaxy systems (stellar mass $1\times10^8<M_* < 5\times10^9\msun$) in a cosmological context.

Massive galaxy ($M_* > 5\times10^9\msun$) mergers and interactions can be observationally identified using methods like asymmetry and concentration measures~\citep{conselice2000, lotz04}. Numerous studies have used these measures to study the merger and interaction rates of massive galaxies as a function of time~\citep[e.g.,][]{lotz11,casteels14}.

However, \citet{martin21} showed that mergers only account for 10\% of morphological disturbances in dwarfs, and that a majority of their morphological disturbances derive from non-merger interactions, such as gas inflows/outflows, close passages, and fly-bys. This means that the typical methods of classifying interactions and mergers for massive galaxies are unlikely to be robust in lower mass regimes. This is particularly true at high redshift, when gas accretion along filaments is expected to be more significant~\citep{keres09}. Thus, it is critical to establish an alternative, cosmologically consistent, way to determine merger and interaction rates for dwarf galaxy pairs from observations.

Another observational technique to determine merger/interaction rates relies upon the pair fraction: the ratio of the number of host galaxies with a companion to the number of potential hosts. Pair fractions for massive galaxies have been measured observationally as a function of time~\citep[e.g.,][]{lotz08,conselice09}, and calibrated against theoretical simulations to determine observability timescales that enable the translation of pair fractions to merger rates \citep{lotz11,kaviraj15,snyder17,qu17,oleary21b}.
In contrast, pair fractions for dwarf primaries are only starting to be measured at high redshift \citep[e.g.][]{ventou17} and have only been studied with small sample size or calibrated against simulations at low redshift \citep{besla18, casteels14,stierwalt15, pearson18}. Thus we are currently missing a theoretical framework to interpret future high redshift observations of dwarf pairs, i.e. with \textit{JWST}. In this study, we quantify the dwarf galaxy pair fraction as a function of time out to $z\sim$4, and compare these results to the massive galaxy pair fraction, using the N-body and hydrodynamic \textit{Illustris-1} cosmological simulations~\citep{vogelsberger14A,nelson15}.

There are examples of galaxy pairs involving dwarf galaxies in our Local Group, including the M31-M33, MW-LMC, and the LMC-SMC (Magellanic Clouds; MCs) systems. The frequency of such pairs have been studied extensively observationally and theoretically at low redshift.
In SDSS DR7, it is found that 11\% of MW/M31-like objects have one massive satellite (LMC/M33 analog) within 150 kpc and 300 km s$^{-1}$~\citep{liu11}. Within 250 kpc, the frequency increases to 42\% (Tollerud 2011). % , robotham12
This is in agreement with $\Lambda$CDM predictions \citep{busha11, bk11, patel17a, gonzalez13}. 
% check gonzalez
Furthermore, 4-6\% of LMC/M33 analogs in both SDSS and cosmological simulations are expected to have a companion at least as massive as an SMC in the field \citep{besla18, sales11}. %% ADD SALES RESULTS ?      
% add how defined here

However, we do not how common these pair configurations are as a function of time in the $\Lambda$CDM framework, leaving open the question as to whether our Local Group is unique in time. 
Recent advances in astrometry have furthermore enabled the measurement of 6D phase space information for these Local Group galaxy pairs.  
In this study, we will examine whether not only the frequency, but also the dynamics of pairs in our Local Group are unique \textit{across cosmic time}.
% HOW DO MASSIVE PAIRS SCALE TO DWARF REGIME

The Local Group also contains a massive major galaxy pair (stellar mass ratio $>$1:4), the MW-M31. For low redshift (z$<$ $\sim$.1) massive galaxies (M$_\ast > 10^{10}$ M$_\odot$), the fraction of major pairs are expected to increase as a function of mass \citep{casteels14}, resulting in major merger rates that rise  with higher mass~\citep{rg15,ventou17,oleary21b}. Additionally, the fraction of major pairs is also measured to increase with increasing redshift~\citep{deravel09,bundy09,duncan19}. In particular \citet{duncan19} find no strong evolution in the relative number of minor to major mergers for massive hosts out to z$<$3. But it is unclear whether these trends for the frequency of major pairs will be the same if the host mass is in the dwarf galaxy regime.

It is important to study the dwarf and massive host regimes separately, as baryonic physics, including gravitational tides, star formation, and outflows, impacts the growth of dwarf galaxies more strongly~\citep{bullockbk17,jeon19,jeon21}. For example, it has been shown that baryonic physics implemented in the \IH\ simulation suppresses the abundance of low mass subhalos ($\leq$10$^{10}\msun$), but such a suppression is not present for massive halos~\citep{chua17}. Thus, the pair fractions and dynamics of dwarf major and minor pairs may change drastically between the dark-matter-only and full-physics simulations, with less differences for massive  pairs. In this study, we illustrate that the impact of baryonic physics on the growth of low mass halos in cosmological simulations may be directly testable with pair fraction observations.

Here we present the first cosmologically motivated constraints on the dwarf pair fraction for major and minor pairs as a function of time. We compare these results to those found for massive major and minor pairs, and address the differences in expected pair fractions between the N-body and full hydrodynamic simulations. This work will furthermore, self-consistently, place known galaxy pairs in our Local Group in a cosmological and temporal context.  In Sec.~\ref{sec:methods}, we introduce the simulations and selection criteria used to perform the analysis of our sample of pairs. In Sec.~\ref{sec:results}, we present the results of this pair fraction study, including preliminary results on the kinematics of dwarf pairs over cosmic time. In Sec.~\ref{sec:discussion}, we discuss the implications of our results on interpretations of future surveys, and what this could mean for constraints of feedback mechanisms in cosmological simulations. In Sec.~\ref{sec:conclusions} we present a summary of the main conclusions of this work, including expected pair fractions for dwarf and massive galaxy pairs.

%%%%%%%%%%%%%%%%%
\section{Sample Selection} \label{sec:methods}
%%%%%%%%%%%%%%%%%
Our sample of dwarf galaxy pairs and massive galaxy pairs from cosmological simulations was chosen following a strict set of selection criteria. In Sec.~\ref{sec:simData} we will detail the simulations from which we collected our data, Sec.~\ref{sec:abundanceMatching} explains the abundance matching prescription used to associate dark matter halos with stellar masses, and Sec.~\ref{sec:selectionCrit} outlines the isolation and mass criteria used to identify our sample.

\subsection{Simulation data}\label{sec:simData}
We utilize data from the \textit{Illustris Project} ~\citep{vogelsberger14A,nelson15}, a set of N-body and hydrodynamic cosmological simulations, to statistically analyse the occurrence rates of dwarf and massive dark matter subhalo pairs over cosmic time. In particular, we are using the highest resolution dark matter only simulation, \textit{Illustris-1-Dark}, and the  full hydrodynamic simulation, \textit{Illustris-1} (hereafter \textit{Illustris-Dark} and \textit{Illustris-Hydro} respectively, see Table~\ref{table:illustris}).
Each of these simulations follows the dynamical evolution of $1820^3$ dark matter particles from $z=127$ to $z=0$ through a periodic volume of $106.5^3$ Mpc$^3$. The Illustris simulations use a cosmology that is consistent with \textit{WMAP-9} ~\citep{hinshaw13}: $\Omega_m = 0.2726$, $\Omega_\Lambda = 0.7274$, $\Omega_b = 0.0456$, $\sigma_8 = 0.809$, $n_s=0.963$, and $h=0.704$.

% Maybe doesn't need whole paragraph but, about dark vs. hydro
We are utilizing data from both \textit{Illustris-Dark} and \textit{Illustris-Hydro} to study the effect of baryonic content on the subhalos. As dwarf subhalos have more shallow dark matter potentials than their more massive counterparts, they are more likely to be subject to disruption caused by baryonic feedback at higher redshifts, which may result in fewer low mass subhalo pairs in \textit{Illustris-Hydro}.

% Subfind paragraph: 
We use catalogs of dark matter halos and their associated subhalos identified in each simulation using the \texttt{SUBFIND} algorithm ~\citep{springel01,dolag09}. Dark matter halos were identified using the Friends-of-Friends (FoF) algorithm ~\citep{davis85}, which links particles that have separations less than $0.2d$, where $d$ is the mean inter-particle separation. Each FoF group, which we will further refer to as a ``group", was further divided into over-dense and gravitationally bound dark matter subhalos.

% Sublink paragraph:
In order to trace the growth and histories of our subhalos, we are using the merger trees provided by the \texttt{SUBLINK} algorithm \citep{gomez15}, which tracks subhalos between snapshots using their particle data. These merger trees provide information about the kinematics and mass growth of subhalos in both \textit{Illustris-Dark} and \textit{Illustris-Hydro}, and in particular will be used to provide the maximal masses of each of the subhalos we investigate.

The lower mass threshold for subhalos that constitute our dwarf pairs have a halo mass greater than $1\times10^9\msun$, such that each of our halos are resolved into at least 130 particles in \textit{Illustris-Dark} and 150 particles in \textit{Illustris-Hydro}. 
% \kc{add explanation for why 100 particles is good enough for the resolution we're looking for.}

\begin{table}[htb]
\begin{tabular}{lll}
 & Illustris-Dark & Illustris-Hydro \\\hline\hline
 L$_{\rm Box}$ [Mpc] & 106.5 & 106.5 \\
m$_{\rm DM}$ [M$_\odot$] & 7.5$\times10^6$ & 6.3$\times10^6$ \\\hline
\end{tabular}
\caption{\label{table:illustris}Parameters of the Illustris simulations used in this analysis.}
\end{table}
% when i start using TNG can add this!
% \begin{table*}[htb]
% \begin{tabular}{lllll}
%  & Illustris-Dark & Illustris-Hydro & IllustrisTNG-Dark & IllustrisTNG-Hydro\\\hline\hline
%  L$_{\rm Box}$ [Mpc] & 106.5 & 106.5 & 110.7 & 110.7 \\
% m$_{\rm DM}$ [M$_\odot$] & 7.5$\times10^6$ & 6.3$\times10^6$ & 8.9$\times10^6$ & $7.5\times10^6$\\\hline
% \end{tabular}
% \caption{Parameters of the Illustris simulations used in this analysis.}
% \end{table*}

\begin{table*}[htb]
  \begin{center}
    \begin{tabular}{lcc}
     & Dwarf Pairs & Massive Pairs \\\hline\hline
    Group Mass & $8\times10^{10} < \rm M_g < 5\times10^{11} M_\odot$ & $1\times10^{12} < \rm M_g < 4\times10^{12} M_\odot$  \\
    % Subhalo Max Mass &  &  \\
    Subhalo Current Mass & $>1\times10^{10} \msun$ & $> 5\times10^{11} \msun$ \\
    Primary &$1\times10^{8} < \rm M_* < 5\times10^{9} \msun$ & $5\times10^{9} < \rm M_* < 1\times10^{11} \msun$\\\hline
    \end{tabular}
    \caption{\label{table:mass}Selection criteria for dwarf and massive pairs.}
  \end{center}
    \end{table*}

\subsection{Abundance matching}\label{sec:abundanceMatching}
Although \textit{Illustris-Hydro} provides stellar masses for all subhalos, however these reported masses are higher than observations suggest~\citep{genel14}. We assign stellar masses to dark matter halos in a consistent way between both \ID\ and \IH using the abundance matching (AM) prescription of \citet{moster13} to associate each dark matter halo with a stellar mass.

The AM relationship at low stellar masses has a large spread, thus a subhalo with a dark matter halo mass of $10^{11}$ may be associated with a stellar mass anywhere between $1.2\times10^{8} - 5\times10^{9}\rm M_\odot$, as can be seen in Fig.~\ref{fig:abundanceMatching}.

There are two inputs to the AM relationship: dark matter mass and redshift. As shown recently by \citet{munshi21}, the relationship between stellar mass and peak halo mass of dwarf subhalos in the Marvel and Justice League simulations is much tighter than the relationship between the stellar mass and halo mass of subhalos at $z=0$. We use the maximal dark matter mass of the halo to ensure that the stellar mass associated with each subhalo reflects the shape of the dark matter profile in the inner regions of the subhalo, thus preventing us from assigning very low stellar masses to halos that have been stripped of %a lot of 
their outer dark matter mass by recent interactions. We evaluate the AM relation using the redshift of the halo at the snapshot that we are interested in, rather than the redshift of the halo at its maximal mass, in order to account for ongoing star formation past peak halo mass.

For the collection of subhalos identified at each snapshot of the simulation, the stellar masses are drawn from the gaussian sampled AM relation. We thus generate 50 realizations for the stellar mass distribution of the same subhalo population in order to account for uncertainties in the AM relation. Each realization is treated as an independent sample of galaxies, where computed properties (stellar masses, relative speeds, etc) are averaged over each realization. We report the mean of all realizations and error bars indicate the 1$\sigma$ dispersion across the realizations.

>>>>

\documentclass[twocolumn]{aastex631}

% Packages
\usepackage{microtype}  % ALWAYS!
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{multirow}
\usepackage{tikz}
\usepackage{xcolor}
\usepackage{soul}
\usepackage{hyperref}

\definecolor{pink}{RGB}{232,132,161}
\definecolor{yellow}{RGB}{255,213,0}

\newcommand{\kc}[1]{\textcolor{yellow}{\textbf{kc: #1}} }
\newcommand\shadetext[2][]{%
  \setbox0=\hbox{{\special{pdf:literal 7 Tr }#2}}%
  \tikz[baseline=0]\path [#1] \pgfextra{\rlap{\copy0}} (0,-\dp0) rectangle (\wd0,\ht0);% 
  }
\newcommand{\gb}[1]{\shadetext[left color=blue, right color=red, middle color=lime, shading angle=45]{\textbf{g: #1}} }
% \newcommand{\ecite}[1]{\textcolor{pink}{\textbf{: #1}} }
% \newcommand{\e}[1]{\textcolor{yellow}{\textbf{: #1}} }

\newcommand{\add}[1]{\textcolor{red}{#1}}

\newcommand{\mlg}{\ensuremath{M_{\rm LG}}}
\newcommand{\mmto}{\ensuremath{M_{\rm M31}}}
\newcommand{\mmw}{\ensuremath{M_{\rm MW}}}
\newcommand{\vtan}{\ensuremath{v_\textrm{tan}}}
\newcommand{\vrad}{\ensuremath{v_\textrm{rad}}}
\newcommand{\ms}[1]{\ensuremath{M_{*{#1}}}}
\newcommand{\reflabel}[1]{\ensuremath{^{\mbox{\scriptsize{#1}}}}}
\newcommand{\scsep}{\ensuremath{\rm r_{sep}/r_{vir}}}
\newcommand{\scvel}{\ensuremath{\rm v_{rel}/v_{vir}}}
\newcommand{\subcat}{\textit{Subhalo Catalog}}
%\newcommand{\lsubcat}{\textit{Low Mass Subhalo Catalog}}
%\newcommand{\hsubcat}{\textit{High Mass Subhalo Catalog}}
\newcommand{\starcat}{\textit{Subhalo + Stellar Mass Catalog}}
\newcommand{\paircat}{\textit{Full Pair Catalog}}
\newcommand{\Rvir}{\ensuremath{\rm R_{vir}}}

% Style tweaks
% \renewcommand{\twocolumngrid}{\onecolumngrid}
% \setlength{\parindent}{1.1\baselineskip}
% \sloppy\sloppypar\raggedbottom\frenchspacing

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\shorttitle{Pair fractions in TNG100}
\shortauthors{Chamberlain et al.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\graphicspath{{./}{../plots/paper1/}}
\input{math_definitions.tex}

% Affiliations
\newcommand{\affuofa}{University of Arizona, 933 N. Cherry Ave, Tucson, AZ 85721, USA}
\newcommand{\affunam}{Instituto de Radioastronom\'ia y Astrof\'isica, Universidad Nacional Aut\'onoma de M\'exico, Apdo. Postal 72-3, 58089 Morelia, Mexico}
\newcommand{\affuoff}{Department of Astronomy, University of Florida, P.O. Box 112055, Gainesville, FL 32611, USA}
\newcommand{\affberk}{Department of Astronomy, University of California, Berkeley, 501 Campbell Hall, Berkeley, CA, 94720, USA}
\newcommand{\affuofu}{Department of Physics and Astronomy, University of Utah, 115 South 1400 East, Salt Lake City, Utah 84112, USA}
\newcommand{\affnyu}{The Center for Cosmology and Particle Physics, New York University, 726 Broadway, New York, NY 10003}
\newcommand{\affnrao}{National Radio Astronomy Observatory, 520 Edgemont Road,
Charlottesville, VA 22903, USA}
\newcommand{\affuva}{Department of Astronomy, University of Virginia, 530 McCormick Road, Charlottesville, VA 22904, USA}
\newcommand{\affempty}{Empty}
\newcommand{\affnottingham}{School of Physics and Astronomy, University of Nottingham, University Park, Nottingham NG7 2RD, UK}
\newcommand{\affkorea}{Korea Astronomy and Space Science Institute, 776 Daedeokdae-ro, Yuseong-gu, Daejeon 34055, Korea}
\newcommand{\afftrent}{Department of Physics and Astronomy, Trent University, 1600 West Bank Drive, Peterborough, ON K9L 0G2, Canada}
\newcommand{\affoccidental}{Occidental College, Physics Department, 1600 Campus Road, Los Angeles, CA 90041, USA}

%% This is the end of the preamble.

\begin{document}

\title{
  % Pair fractions of low mass and high mass galaxy pairs in the TNG100 cosmological simulation
  % low mass and high mass galaxy pairs evolve differently as a function of redshift
  %A physically motivated framework for measuring\\ the mass and redshift dependence of galaxy pair fractions across cosmic time 
  \add{A physically motivated framework to compare pair fractions\\ of isolated low and high mass galaxies across cosmic time }

   % for measuring galaxy pair fractions across cosmic time
   % measuring or recovering?
  }
% Differences in the redshift evolution of low mass and high mass galaxy pairs in the TNG100 cosmological simulation

\author[0000-0001-8765-8670]{Katie~Chamberlain}
\thanks{E-mail: katiechambe@arizona.edu}\affiliation{\affuofa}

\author[0000-0003-0715-2173]{Gurtina Besla}
\affiliation{\affuofa}

\author[0000-0002-9820-1219]{Ekta~Patel}
\thanks{Hubble Fellow}\affiliation{\affberk} \affiliation{\affuofu}

\author[0000-0002-9495-0079]{Vicente Rodriguez-Gomez}
\affiliation{\affunam}

\author[0000-0002-5653-0786]{Paul Torrey}
\affiliation{\affuoff}

\author[0000-0003-2939-8668]{Garreth Martin}
\affiliation{\affnottingham}\affiliation{\affkorea}\affiliation{\affuofa}

\author[0000-0001-8348-2671]{Kelsey Johnson}
\affiliation{\affuva}

\author[0000-0002-3204-1742]{Nitya Kallivayalil}
\affiliation{\affuva}

\author[0000-0002-1871-4154]{David Patton}
\affiliation{\afftrent}

\author[0000-0003-0256-5446]{Sarah Pearson}
\thanks{Hubble Fellow}\affiliation{\affnyu}

\author[0000-0003-3474-1125]{George Privon}
\affiliation{\affnrao}\affiliation{\affuoff}\affiliation{\affuva}

\author[0000-0002-2596-8531]{Sabrina Stierwalt}
\affiliation{\affoccidental}

\begin{abstract} 
    Low mass galaxy pair fractions are under-studied across cosmic time. \add{Further, it is unclear whether low mass pair fractions evolve in the same way as more massive systems over cosmic time.}
    \add{In the era of JWST, Roman, and Rubin, selecting galaxy pairs in a self-consistent way will be critical to connect observed pair fractions to cosmological merger rates across all mass scales and redshifts.}
    \add{We use the Illustris TNG100 simulation to create a sample of physically associated pairs between $z=0-4.2$, including low mass ($\rm 10^8<M_*<5\times10^9\,\Msun$) and high mass ($\rm 5\times10^9<M_*<10^{11}\,\Msun$) isolated subhalo pairs.}
    The low mass pair fraction, i.e. the fraction of galaxies in pairs, increases from $z=0-2.5$, while the high mass pair fraction peaks at $z=0$ and is constant or slightly decreasing at $z>1$. 
    At $z=0$ the low mass major (1:4 mass ratio) pair fraction is 4$\times$ lower than high mass pairs, consistent with findings for cosmological merger rates. 
    \add{We show that separation limits that vary with the mass and redshift of the system, such as scaling by the virial radius of the host halo ($r_{\mathrm{sep}}< 1 R_{\rm vir}$), are critical for recovering pair fraction differences between low mass and high mass systems.}
    \add{Alternatively, static physical separation limits applied equivalently to all galaxy pairs do not recover the differences between low and high mass pair fractions, even up to separations of $300\,\kpc$.} 
    Finally, we place isolated mass-analogs of Local Group galaxy pairs (i.e., MW--M31, MW--LMC, LMC--SMC) in a cosmological context, showing that isolated analogs of LMC--SMC-mass pairs, and low separation ($<50\kpc$) MW--LMC-mass pairs, are $2-3\times$ more common at $z\gtrsim2-3$.

    % old abstract:
    % Low mass galaxy pair fractions are under-studied across cosmic time. 
    % In the era of JWST, Roman, and Rubin, a self-consistent framework is needed to select both low and high mass galaxy pairs to connect observed pair fractions to cosmological merger rates across all mass scales and redshifts. We use the Illustris TNG100 simulation to identify physically associated pairs between $z=0-4.2$. 
    % Our sample includes low mass ($\rm 10^8<M_*<5\times10^9\,\Msun$) and high mass ($\rm 5\times10^9<M_*<10^{11}\,\Msun$) isolated subhalo pairs, with stellar masses from abundance matching. 
    % The low mass pair fraction, i.e. the fraction of galaxies in pairs, increases from $z=0-2.5$, while the high mass pair fraction peaks at $z=0$ and is constant or slightly decreasing at $z>1$. 
    % At $z=0$ the low mass major (1:4 mass ratio) pair fraction is 4$\times$ lower than high mass pairs, consistent with findings for cosmological merger rates. 
    % Our results indicate that pair fractions can faithfully reproduce trends in merger rates if galaxy pairs are selected appropriately. 
    % Specifically, static pair separation limits applied equivalently to all galaxy pairs do not recover the evolution of low and high mass pair fractions. 
    % Instead, we advocate for separation limits that vary with the mass and redshift of the system, such as separation limits scaled by the virial radius of the host halo ($r_{\mathrm{sep}}< 1 R_{\rm vir}$). 
    % Finally, we place isolated mass-analogs of Local Group galaxy pairs (i.e., MW--M31, MW--LMC, LMC--SMC) in a cosmological context, showing that isolated analogs of LMC--SMC-mass pairs, and low separation ($<50\kpc$) MW--LMC-mass pairs, are $2-3\times$ more common at $z\gtrsim2-3$.
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction} \label{sec:intro}
    Galaxy mergers have been studied in detail as a mechanism for driving galaxy evolution, and have been identified as a trigger of, for example, 
    active galactic nuclei (AGN)~\citep[e.g.][]{Hopkins2008,Treister2010,Ramos2011,Satyapal2014,Comerford2015,Glikman2015,Blecha2018,Ellison2019}, 
    star formation~\citep[e.g.][]{Mihos1996, DiMatteo2008,Ellison2011,Patton2013, Hopkins2013,Martin2017,Patton2020,Hani2020,Martin2021}, 
    and changes in morphology~\citep[e.g.][]{Conselice2003,Lotz2008,Casteels2014,Patton2016,Bignone2017,RG2017,Martin2018,Jackson2019,Snyder2019,Jackson2022,Guzman-Ortega2023}. 
    Constraining the merger rate for galaxies is critical for quantifying the importance of mergers for galaxy evolution and testing predictions for hierarchical assembly in cold dark matter theory \citep[e.g.][]{Stewart2009,Hopkins2010,RG2015}.

    In practice, merger rates cannot be measured straightforwardly from observations and, rather, are calculated by converting the observed frequency of a galaxy merger signature (i.e., asymmetry, concentration, pair fraction, etc.) to a merger rate using an observability timescale~\citep[e.g.][]{Lotz2011}. However, morphological signatures of mergers are often caused by non-merger phenomena.
    For example, low mass galaxies are more commonly morphologically disturbed by fly-bys and non-merger interactions than by mergers~\citep{Martin2021}, 
    and star-forming galaxies at high redshift tend to be clumpy and asymmetric even in isolation~\citep{Wuyts2013}.
    Additionally, detection of these signatures can be strongly dependent on image depth or galaxy stellar mass, and identifying tidal features that rely on human classifications may be unreliable~\citep{Martin2022}.

    Pair fractions, on the other hand, can be calculated independently of morphological information, and thus offer a more robust method of observationally quantifying merger rates across cosmic time. 
    Pair fractions of high mass ($M_*\gtrsim10^{10}\Msun$) galaxies have been well studied across cosmic time both observationally~\citep[e.g.][]{Patton2002,Lin2004,Lin2008,Lotz2011,Ferreras2014, Man2016,Duncan2019}, and theoretically~\citep[e.g.][]{RG2015,Snyder2017,Snyder2023}, with theoretical studies typically done in projection for comparison to observational campaigns\footnote{The studies cited here have limits on the projected separation of their sample, such that the projected separations range between $5\,\kpc$ and $\sim140\,\kpc$.}.  
    Low mass galaxy pairs ($10^8<M_*<5\times10^9\Msun$) have been studied at low redshift~\citep[e.g.][]{Stierwalt2015,Pearson2016,Besla2018,Paudel2018,Luber2022},
    but are less well understood across cosmic time owing to the difficulty in observing faint systems outside of the Local Volume. 
    However, JWST~\citep{Gardner2006}, as well as the next generation of deep and wide-field surveys from Rubin Observatory\footnote{ \href{https://rubinobservatory.org/}{https://rubinobservatory.org}}~\citep{RUBIN2019} and Roman Space Telescope\footnote{\href{https://roman.gsfc.nasa.gov/}{https://roman.gsfc.nasa.gov}}~\citep{Spergel2015}, will significantly revolutionize our ability to identify such systems at high redshift $(z\gtrsim10)$~\citep{Behroozi2020} and in abundance at lower redshift ($z\lesssim6$)~\citep{Robertson2019a,Robertson2019b-LSST}.

    There are reasons to believe that low mass and high mass pairs evolve differently as a function of time.
    For example, semi-empirical and cosmological studies find that galaxy merger rates vary both with redshift and the mass of the most massive galaxy of the pair~\citep[see e.g.][]{Guo2008,Stewart2009,Hopkins2010,RG2015,Martin2021}.
    % Such differences are likely to give rise to differences in the pair fractions of low and high mass pairs.
    It is thus reasonable to assume that the pair fractions of these two mass scales reflect these evolutionary differences as well.

    The redshift evolution of low and high mass pair fractions has not yet been studied in simulations in a self-consistent way, where high mass and low mass pairs are selected from simulations using otherwise equivalent selection criteria. 
    We aim to characterize the redshift behavior of pair fractions of low mass and high mass pairs, independent of environmental and projection effects, and to create a robust framework in which to fairly compare pair fractions of different mass scales across cosmic time.

    In particular, we take the approach of consistently selecting physically associated pairs of low mass ($10^8<M_*<5\times10^9\Msun$) and high mass $(5\times10^9<M_*<10^{11}\,\Msun)$ galaxies. 
    Specifically, we identify major ($1-1/4$) and minor $(1/10-1/4)$ stellar mass ratio pairs from $z=0-4.2$ in the IllustrisTNG cosmological simulation, TNG100. 
    We require that pairs are part of the same Friends of Friends (FoF) group and that no other more massive perturbers are nearby, which allows us to ensure that the recovered pair fractions are inherent to the population of selected pairs rather than a function of environment.

    Typically, 
    \add{pair fraction studies via simulations and observations}
    apply physical projected separation cuts that are constant over time and do not vary with the mass of the target system. 
    \add{In this study we identify how the application of a fixed separation criteria affects inferred pair fractions, and show that a time and mass evolving separation cut is necessary to permit equitable pair fraction comparisons across different mass regimes.} 
    \add{Although implementing alternative selection criteria in future observational pair fraction studies may not always be strictly necessary, we recommend that pair selection criteria be reevaluated for observational studies that seek to compare pair fractions as a function of mass or redshift.}
    % In this study we identify how the application of a fixed separation criteria affects inferred pair fractions, and make recommendations for a time and mass evolving separation cut to permit equitable pair fraction comparisons between different mass regimes.
    % To fairly test for differences in the evolutionary behavior of pair fractions, especially in observational studies, it is necessary to determine whether this is an appropriate technique for accurately recovering true pair fraction behavior.

    Finally, there are a number of galaxy pairs in the Local Group that are mass analogs of the isolated pairs in our sample.
    For example: the MW--M31 system is a high mass major pair, the MW--LMC and M31--M33 systems are high mass minor pairs, and the LMC--SMC are a low mass minor pair.
    Studies have examined the frequency of such configurations, particularly at low redshift or in the form of progenitor systems of the Local Group in cosmological simulations~\citep{Bk2011,Fattahi2013,Patel2017a-Orbits,Geha2017,Mao2021}, but the prevalence of such pairs in isolation has not yet been quantified as a function of redshift. 
    We utilize our pair dataset to quantify the likelihood of finding isolated mass-analogs of Local Group pairs as a function of redshift, particularly when the present day separations of these systems are folded in.

    This paper is structured as follows. 
    In Section~\ref{sec:methods}, we outline our methodology for selecting physically associated high mass and low mass pairs from the TNG simulation. 
    In Section~\ref{sec:pairprops}, we provide an overview of the properties of the selected sample, including the number of primaries, pairs, and their stellar mass ratios as a function of time. 
    We present the time-evolving pair fraction of high mass and low mass pairs in Section~\ref{sec:results}, and show how they change for different separation criteria. 
    \add{In Section~\ref{sec:discussion}, we give context to our results by drawing comparisons to Local Group pairs and other Illustris-based pair fraction studies, and discuss implications for observational campaigns.}
        % In Section~\ref{sec:discussion}, we discuss the implications of the differences between low and high mass pair fractions for observational campaigns and give context to our results by drawing comparisons to Local Group pairs. 
    Finally, we summarize our results and conclusions in Section~\ref{sec:summary}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
